---
title: "Data Appendix"
author: "YIXIAO XIE and YEYUXI YI"
date: "December 14, 2023"
output: pdf_document
professor: Dr. Michael O'Hara
header-includes: \usepackage{setspace}\doublespacing
course: ECON 342 01 Econometrics
---
```{r include = FALSE, echo = FALSE}

rm(list = ls())


library(dplyr)
library(ggplot2)
library(stargazer)
library(readr)
library(dagitty)
library(ggdag)
library(car)
library(knitr)
```

```{r echo = FALSE}
StockNewsData <- read.csv2("StockNewsData.csv", header = TRUE, na.strings="")
```

The following data summarizes the properties of the collected data and variables. All data is from either the Bloomberg or Yahoo finance. There is total 161 observation. 


```{r echo = FALSE}
stargazer(as.data.frame(StockNewsData), type = "text", header = FALSE, keep.stat = c("n", "rsq"), title = "Table 1: Summary Statistics")
```


D1~20: Return of a stock on Day 1~20 of earning report release
M1~20: Return of the market on Day 1~20 of earning report release
Beta:the linear relationship of the stock related to the market based on historical data
Dnews – News Categories --- Extremely Bad/Bad/No News/Good/Extremely Good

Now we want to see the stock return's normal distribution.
```{r echo = FALSE}
ggplot(StockNewsData,aes(x = D1))+
  geom_histogram()

ggplot(StockNewsData,aes(x = D2))+
  geom_histogram()

ggplot(StockNewsData,aes(x = D3))+
  geom_histogram()

ggplot(StockNewsData,aes(x = D5))+
  geom_histogram()

ggplot(StockNewsData,aes(x = D10))+
  geom_histogram()

ggplot(StockNewsData,aes(x = D20))+
  geom_histogram()

```


Now we want to see the market return's normal distribution
```{r echo = FALSE}
ggplot(StockNewsData,aes(x = M1*Beta))+
  geom_histogram()

ggplot(StockNewsData,aes(x = M2*Beta))+
  geom_histogram()

ggplot(StockNewsData,aes(x = M3*Beta))+
  geom_histogram()

ggplot(StockNewsData,aes(x = M5*Beta))+
  geom_histogram()

ggplot(StockNewsData,aes(x = M10*Beta))+
  geom_histogram()

ggplot(StockNewsData,aes(x = M20*Beta))+
  geom_histogram()
```

Under different news categories whether the stock return is distributed or not
```{r echo = FALSE}
ggplot(StockNewsData,aes(y = Dnews,x = D1))+
  geom_point()

ggplot(StockNewsData,aes(y = Dnews,x = D2))+
  geom_point()

ggplot(StockNewsData,aes(y = Dnews,x = D3))+
  geom_point()

ggplot(StockNewsData,aes(y = Dnews,x = D5))+
  geom_point()

ggplot(StockNewsData,aes(y = Dnews,x = D10))+
  geom_point()

ggplot(StockNewsData,aes(y = Dnews,x = D20))+
  geom_point()
```

Whether there is a linear relationship between stock return and market return. 

```{r echo = FALSE}
ggplot(StockNewsData,aes(y = D1,x = M1*Beta))+
  geom_point()
ggplot(StockNewsData,aes(y = D2,x = M2*Beta))+
  geom_point()
ggplot(StockNewsData,aes(y = D3,x = M3*Beta))+
  geom_point()
ggplot(StockNewsData,aes(y = D5,x = M5*Beta))+
  geom_point()
ggplot(StockNewsData,aes(y = D10,x = M10*Beta))+
  geom_point()
ggplot(StockNewsData,aes(y = D20,x = M20*Beta))+
  geom_point()
```













